BOLD Parameter Estimation Using Particle Filters

### 2010-07-18 doc/thesis/results.tex
"done" with single timeseries simulations, on the the real results next then back to the slice sim

\subsection{Signal with Low Noise}
%tests had noise of $ \{\sigma_y = .01, \sigma_x = .005\} $, where $\sigma_y$ is the
%measurement noise, and $\sigma_x$ is the wiener step size. Both signals used the
%parameters .
%The particle filter used the parameters defined in \autoref{tab:Prior} (\autoref{sec:PriorDistrib}),
%thus the particle filter was not centered over the correct values. 
\begin{figure}
\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/realization_lownoise}
\caption{Test Signals with low noise compared to the clean signal.}
\label{fig:LowNoiseRealization}
\end{figure}

@@

\begin{table}[t]
\centering
\begin{tabular}{|c | c | c | c | c | c | c | c | c | c |}
\hline 
$\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$  & $ \sum \tau $ & $\sqrt{MSR}$ &$\sqrt{MSE}$\\
\hline 
\rowcolor[gray]{.8}
1.45 & .3 & .47 & .044 & 1.94 & 1.99 & 1.8  & 5.38 &  & \\
\hline 
\hline 
1.2221 & 0.3449 & 0.3346 & 0.0714 & 1.6045 & 2.2753 & 1.5945 & 5.1019 &  0.003211  & 0.009876  \\
1.3749 & 0.3318 & 0.3630 & 0.0733 & 1.6408 & 2.1030 & 1.5763 & 5.1187 &  0.003055  & 0.009932  \\
1.1660 & 0.3221 & 0.3406 & 0.0822 & 1.6477 & 2.3535 & 1.2452 & 5.1672 &  0.003289  & 0.009680  \\
1.2318 & 0.3271 & 0.3403 & 0.0796 & 1.6270 & 2.1852 & 1.3033 & 5.0439 &  0.002847  & 0.009120  \\
1.1832 & 0.3179 & 0.3472 & 0.0821 & 1.5496 & 2.2912 & 1.2782 & 5.0240 &  0.003006  & 0.009713  \\
1.1424 & 0.334  & 0.3473 & 0.0737 & 1.6221 & 2.2908 & 1.4025 & 5.0553 &  0.002833  & 0.009485  \\
1.3004 & 0.3596 & 0.3564 & 0.0768 & 1.5641 & 2.1323 & 1.6034 & 4.9968 &  0.003028  & 0.010219  \\
1.2401 & 0.3460 & 0.3398 & 0.0891 & 1.6499 & 2.2366 & 1.2900 & 5.1265 &  0.003044  & 0.010080  \\
1.1709 & 0.3274 & 0.3464 & 0.0826 & 1.5373 & 2.2826 & 1.3783 & 4.9909 &  0.003345  & 0.010329  \\
1.1897 & 0.3434 & 0.3355 & 0.0798 & 1.5358 & 2.3075 & 1.4277 & 5.0330 &  0.003175  & 0.010015  \\
1.184 &  0.3405 & 0.3502 & 0.0892 & 1.6103 & 2.2793 & 1.1645 & 5.0735 &  0.002889  & 0.009505  \\
\hline                                                                           
1.2187 & 0.3359 & 0.3456 & 0.0800 & 1.599 & 2.2488 & 1.3876 & 5.0665 & 0.003066     & 0.009814 \\
\hline 
\end{tabular}
\caption{Estimated Parameters on 11 different runs with low noise. First row is the true parameters,
last is mean over the 10 runs. The $\sqrt{MSR}$ is the square root of the mean square of the
residuals. Although in real data I won't distinguish between residual and error, here I will.
$\sqrt{MSE}$ is the mean squared difference between the true signal and the estimated signal.
This is the actual error, due to a combination of noise and preprocessing. }
\label{tab:LowNoiseResults} 
\end{table}

@@

\subsection{High Noise}
\label{sec:HighNoise}
\begin{figure}
\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/realization_highnoise}
\caption{Test Signals with high noise compared to the clean signal, $\sigma_x = .01, \sigma_y=.005$}
\label{fig:HighNoiseRealization}
\end{figure}
\begin{figure}
\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/preprocessed_highnoise}
\caption{A comparison of the preprocessed signals for the high noise case.}
\label{fig:PreprocessedHighNoise}
\end{figure}
\begin{figure}
\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/comparison_highnoise}
\caption{A comparison of the fitted signals for the high noise case.}
\label{fig:FitComparisonHighNoise}
\end{figure}
\begin{figure}
\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/highnoise_56_noise}
\caption{Two particular preprocessed noise realizations for the high noise case.}
\label{fig:NoiseComparisonJustTwo}
\end{figure}

@@

\begin{figure}
\subfigure[$\tau_0$, $\alpha$, $E_0$, $V_0$, Run 1]
{\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=16cm]{images/highnoise_run5_1}}\\
\end{figure}
\begin{figure}
\subfigure[$\tau_s$, $\tau_f$, $\epsilon$, $V$, Run 1] 
{\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=16cm]{images/highnoise_run5_2}}\\
\end{figure}
\begin{figure}
\subfigure[$Q$, $S$, $F$, $BOLD$, Run 1 ]
{\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=16cm]{images/highnoise_run5_3}}
\caption{Converging histogram for parameters during run 1, as in \autoref{fig:NoiseComparisonJustTwo}.}
\label{fig:ConvergenceRuns1}
\end{figure}

\begin{figure}
\subfigure[$\tau_0$, $\alpha$, $E_0$, $V_0$, Run 2]
{\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=16cm]{images/highnoise_run6_1}}\\
\end{figure}
\begin{figure}
\subfigure[$\tau_s$, $\tau_f$, $\epsilon$, $V$, Run 2] 
{\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=16cm]{images/highnoise_run6_2}}\\
\end{figure}
\begin{figure}
\subfigure[$Q$, $S$, $F$, $BOLD$, Run 2 ]
{\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=16cm]{images/highnoise_run6_3}}
\caption{Converging histogram for parameters during run 2, as in \autoref{fig:NoiseComparisonJustTwo}.}
\label{fig:ConvergenceRuns2}
\end{figure}

@@

\subsection{Pure-Noise, low magnitude}
The two final single-voxel tests force the particle filter to attempt to "learn" a noise-only
time series. In the first test the noise used will be the same as that from the \autoref{sec:HighNoise},
$\sigma_x = .01, \sigma_y = .005$. The parameters will be set the same as well, but the
stimulus/input will be set to 0 the entire time. In effect it is a region of the brain for
which no visual stimulus induces activity. For the particle filter, on the other hand,
the stimulus will be left as in the previous two sections. The question is if the particle
filter is able to appropriately differentiate from a signal with noise. The noise time series'
are shown in \autoref{fig:NoiseOnly}, the preprocessed versions are shown in \autoref{fig:NoiseOnlyPrep}.
\begin{figure}[H]
\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm,width=16cm]{images/realization_noiseonly}
\caption{Time-series lacking any real signal. With, $\sigma_x = .01, \sigma_y=.005$}
\label{fig:NoiseOnly}
\end{figure}
\begin{figure}[H]
\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm,width=16cm]{images/preprocessed_noiseonly}
\caption{A comparison of the preprocessed signals for the signal-free case.}
\label{fig:PreprocessedNoiseOnly}
\end{figure}

\begin{table}[t]
\centering
\begin{tabular}{|c | c | c | c | c | c | c | c | c |}
\hline 
$\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$  &  $\sqrt{MSE}$   \\
\hline 
1.0324 & 0.33211 & 0.34058 & 0.03012 & 1.40665 & 2.52079 & 0.5311 & \\
 0.98189 & 0.33047 & 0.3386 & 0.03014 & 1.45707 & 2.47232 & 0.45049 & 0.00158551 \\
 1.0429 & 0.33224 & 0.34124 & 0.02946 & 1.4618 & 2.49245 & 0.43012 &  0.00165055 \\
 1.02054 & 0.3321 & 0.33484 & 0.02586 & 1.45848 & 2.48741 & 0.4193 &  0.00150604 \\
 1.0565 & 0.33405 & 0.33758 & 0.02791 & 1.43784 & 2.52545 & 0.47517 & 0.00151513 \\
 1.01867 & 0.33528 & 0.33918 & 0.02782 & 1.48345 & 2.49605 & 0.44209 &0.00155645 \\
 1.051 & 0.33038 & 0.33837 & 0.02985 & 1.47651 & 2.48621 & 0.42719 &  0.00158526 \\
 1.00281 & 0.32929 & 0.33988 & 0.0298 & 1.43519 & 2.49256 & 0.48899 & 0.00164299 \\
 1.00893 & 0.33273 & 0.33982 & 0.0289 & 1.42903 & 2.49754 & 0.45688 & 0.0016793  \\
 1.01289 & 0.33275 & 0.3376 & 0.02997 & 1.41188 & 2.49881 & 0.50628 & 0.00182622 \\
 1.10247 & 0.33371 & 0.3419 & 0.02939 & 1.43774 & 2.53384 & 0.44079 & 0.00195009 \\
\hline                                                                
1.03009 & 0.33228 & 0.33905 & 0.02902 & 1.44506 & 2.50031 & 0.46076 & 0.00165129 \\
\hline 
\end{tabular}
\caption{Estimated Parameters on 10 different runs with low noise. }
\label{tab:NoiseOnlyResults} 
\end{table}

The line fits are shown in \autoref{fig:fits_noiseonly}. Now the
low MSE and reasonableness of the parameters is cause for concern. Additionally, the precision
with which the convergence occurred is enviable, compared to the actual signals.
However, the impulse response, which has less importance in nonlinear systems, could
be usefull for separating out good voxels from bad. Note that in \autoref{fig:PreprocessedNoiseOnly}
peaks at well below $.005$, the standard deviation of the weighting function. Thus, the reason
the fits are in such agreement is that no convergence occurred at all, as 
further shown by \autoref{JustNoiseConvergence}. Thus one more test is necessary on
a signal with enough noise to actually be considered a signal. But first,
there is still the matter of identifying this noise-only voxel. Clearly MSE and
the parameters themselves don't have the information to differentiate in this case.
Thus the answer falls on either the final estimated variance (which will be high
because no particles strayed far from the center) or the impulse response.
I prefer the impulse response because it doesn't depend on something that 
may be less than dependable, namelay the rate of convergence.

\begin{figure}[H]
\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm,width=16cm]{images/fits_noiseonly}
\caption{Fits to the non-active, low noise signal. Note that the line is thick because all
the fits overlap. This is all 1 fitted lines.}
\label{fig:fits_noiseonly}
\end{figure}

\begin{figure}[H]
\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm,width=16cm]{images/justnoise_fit_0}
\caption{Fit from a single particle filter run, with the noise input. }
\label{fig:justnoise_fit_0}
\end{figure} %uses allnoise/ALLNOISE-0-w0

\begin{figure}[H]
\subfigure[$\tau_0$, $\alpha$, $E_0$, $V_0$]
{\includegraphics[clip=true,trim=7cm 3cm 7cm 4cm, width=15cm]{images/justnoise_hist_1}}\\
\end{figure}
\begin{figure}[H]
\subfigure[$\tau_s$, $\tau_f$, $\epsilon$, $V$] 
{\includegraphics[clip=true,trim=7cm 3cm 7cm 1cm, width=15cm]{images/justnoise_hist_2}}\\
\end{figure}
\begin{figure}[H]
\subfigure[$Q$, $S$, $F$, $BOLD$ ]
{\includegraphics[clip=true,trim=7cm 3cm 7cm 2cm, width=15cm]{images/justnoise_hist_3}}
\caption{Converging histogram for parameters when the signal consists purely of low level noise. 
Same run as \autoref{fig:justnoise_fit_0}}
\label{fig:JustNoiseConvergence}
\end{figure}

To give a frame of reference to the impulse response level, the impulse response of every
parameter estimated, includeing ones from \autoref{sec:BigNoise}, are shown in \autoref{tab:ImpulseResponses}.

\subsection{Pure Noise, High Magnitude}
\label{sec:PureNoiseHighMag}

\begin{figure}[H]
\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm,width=16cm]{images/fits_noiseonly_high}
\caption{Fits to the non-active, high noise signal. Note that the line is thick because all
the fits overlap. This is all 11 fitted lines.}
\label{fig:fits_noiseonly_high}
\end{figure}

\begin{figure}[H]
\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm,width=16cm]{images/justbignoise_fit_0}
\caption{Fit from a single particle filter run, with the noise input. }
\label{fig:justbignoise_fit_0}
\end{figure} %uses allnoise/ALLNOISE-0-w0

\begin{figure}[H]
\subfigure[$\tau_0$, $\alpha$, $E_0$, $V_0$]
{\includegraphics[clip=true,trim=7cm 3cm 7cm 4cm, width=15cm]{images/justbignoise_1}}\\
\end{figure}
\begin{figure}[H]
\subfigure[$\tau_s$, $\tau_f$, $\epsilon$, $V$] 
{\includegraphics[clip=true,trim=7cm 3cm 7cm 1cm, width=15cm]{images/justbignoise_2}}\\
\end{figure}
\begin{figure}[H]
\subfigure[$Q$, $S$, $F$, $BOLD$ ]
{\includegraphics[clip=true,trim=7cm 3cm 7cm 2cm, width=15cm]{images/justbignoise_3}}
\caption{Converging histogram for parameters when the signal consists purely of noise, with peaks comparable  to 
the convential BOLD signal. ($\sigma_y = .01, \sigma_x = .005$). Same run as fit in \autoref{fig:justbignoise_fit_0}}
\label{fig:JustNoiseConvergenceLarge}
\end{figure}

\subsection{Single Voxel Review}
\begin{table}[t]
\centering
\begin{tabular}{|c | c | c | c | c | c | c | c | c |}
\hline 
& \multicolumn{4}{|c|}{Signal} & \multicolumn{4}{|c|}{No Signal}\\
\hline
& \multicolumn{2}{|c|}{Low Noise} & \multicolumn{2}{|c|}{High Noise} 
& \multicolumn{2}{|c|}{$\sigma_y = .001, \sigma_x = .0005$} 
& \multicolumn{2}{|c|}{$\sigma_y = .01, \sigma_x = .005$}\\
\hline 
&$\sqrt{MSE}$ & $N\sqrt{MSE}$ &
 $\sqrt{MSE}$ & $N\sqrt{MSE}$ &
 $\sqrt{MSE}$ & $N\sqrt{MSE}$ &
 $\sqrt{MSE}$ & $N\sqrt{MSE}$ \\
\hline       
\hline       
1 & 0.003211 & 0.478015  & 0.014065  & 1.038941    &0.001667 & 1.295007 & 0.015114 & 1.33641 \\
2 & 0.003055 & 0.531771  & 0.013731  & 0.951648  &  0.001586 & 1.301750 & 0.015510 & 1.336667 \\
3 & 0.003289 & 0.545798  & 0.012754 &  0.995392  &  0.001651 & 1.26287 &  0.013957 & 1.159567 \\
4 & 0.002847 & 0.498237  & 0.016726  & 1.161286  &  0.001506 & 1.431959 & 0.013814 & 1.099876 \\
5 & 0.003006 & 0.468048  & 0.013698 &  1.039724   & 0.001515 & 1.256645 & 0.014350 & 1.201072 \\
6 & 0.002833 & 0.459001  & 0.011495 &  1.002143  &  0.001557 & 1.270797 & 0.013504 & 1.045886 \\
7 & 0.003028 & 0.460962  & 0.015550 &  1.088472   & 0.001585 & 1.154406 & 0.013006 & 1.205429 \\
8 & 0.003044 & 0.518376  & 0.012054 &  1.059620  &  0.001643 & 1.274564 & 0.015446 & 1.122502 \\
9 & 0.003345 & 0.525303  & 0.015104 &  1.015703  &  0.001679 & 1.320244 & 0.014847 & 1.086366 \\
10 & 0.003175 &0.492111  & 0.012493  & 1.189964  &  0.001826 & 1.344562 & 0.013125 & 1.221350 \\
11 & 0.002889 &0.490919  & 0.012165 &  0.953676  &  0.001950 & 1.325223 & 0.014457 & 1.117367 \\
\hline
mean & 0.003066 & 0.497140  & 0.013621 &  1.04514   &  0.001651 & 1.294366 & 0.014285 & 1.175681 \\
\hline
\end{tabular}
\caption{$\sqrt{MSE}$ and the normalized $\sqrt{MSE}$, $\left(\sqrt{MSE}/(\text{Median Absolute Deviation})\right)$  
for different configurations.} \label{tab:SignleVoxelActivationComparison} 
\end{table}

>>>>

%LOW NOISE SECTION, with a signal
\subsection{Signal with Low Noise}
%tests had noise of $ \{\sigma_y = .01, \sigma_x = .005\} $, where $\sigma_y$ is the
%measurement noise, and $\sigma_x$ is the wiener step size. Both signals used the
%parameters .
%The particle filter used the parameters defined in \autoref{tab:Prior} (\autoref{sec:PriorDistrib}),
%thus the particle filter was not centered over the correct values. 
\begin{figure}
\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/realization_lownoise}
\caption{Test Signals with low noise compared to the clean signal.}
\label{fig:LowNoiseRealization}
\end{figure}

@@

\begin{table}[t]
\centering
\begin{tabular}{|c | c | c | c | c | c | c | c | c | c |}
\hline 
$\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$  & $ \sum \tau $ & $\sqrt{MSR}$ &$\sqrt{MSE}$\\
\hline 
\rowcolor[gray]{.8}
1.45 & .3 & .47 & .044 & 1.94 & 1.99 & 1.8  & 5.38 &  & \\
\hline 
\hline 
1.2221 & 0.3449 & 0.3346 & 0.0714 & 1.6045 & 2.2753 & 1.5945 & 5.1019 &  0.003211  & 0.009876  \\
1.3749 & 0.3318 & 0.3630 & 0.0733 & 1.6408 & 2.1030 & 1.5763 & 5.1187 &  0.003055  & 0.009932  \\
1.1660 & 0.3221 & 0.3406 & 0.0822 & 1.6477 & 2.3535 & 1.2452 & 5.1672 &  0.003289  & 0.009680  \\
1.2318 & 0.3271 & 0.3403 & 0.0796 & 1.6270 & 2.1852 & 1.3033 & 5.0439 &  0.002847  & 0.009120  \\
1.1832 & 0.3179 & 0.3472 & 0.0821 & 1.5496 & 2.2912 & 1.2782 & 5.0240 &  0.003006  & 0.009713  \\
1.1424 & 0.334  & 0.3473 & 0.0737 & 1.6221 & 2.2908 & 1.4025 & 5.0553 &  0.002833  & 0.009485  \\
1.3004 & 0.3596 & 0.3564 & 0.0768 & 1.5641 & 2.1323 & 1.6034 & 4.9968 &  0.003028  & 0.010219  \\
1.2401 & 0.3460 & 0.3398 & 0.0891 & 1.6499 & 2.2366 & 1.2900 & 5.1265 &  0.003044  & 0.010080  \\
1.1709 & 0.3274 & 0.3464 & 0.0826 & 1.5373 & 2.2826 & 1.3783 & 4.9909 &  0.003345  & 0.010329  \\
1.1897 & 0.3434 & 0.3355 & 0.0798 & 1.5358 & 2.3075 & 1.4277 & 5.0330 &  0.003175  & 0.010015  \\
1.184 &  0.3405 & 0.3502 & 0.0892 & 1.6103 & 2.2793 & 1.1645 & 5.0735 &  0.002889  & 0.009505  \\
\hline                                                                           
1.2187 & 0.3359 & 0.3456 & 0.0800 & 1.599 & 2.2488 & 1.3876 & 5.0665 & 0.003066     & 0.009814 \\
\hline 
\end{tabular}
\caption{Estimated Parameters on 11 different runs with low noise. First row is the true parameters,
last is mean over the 11 runs. The $\sqrt{MSR}$ is the square root of the mean square of the
residuals. Although in real data I won't distinguish between residual and error, here I will.
$\sqrt{MSE}$ is the mean squared difference between the true signal and the estimated signal.
This is the actual error, due to a combination of noise and preprocessing. }
\label{tab:LowNoiseResults} 
\end{table}

@@

%HIGH NOISE SECTION, with a signal
\subsection{High Noise}
\label{sec:HighNoise}
\begin{figure}
\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/realization_highnoise}
\caption{Test Signals with high noise compared to the clean signal, $\sigma_x = .01, \sigma_y=.005$}
\label{fig:HighNoiseRealization}
\end{figure}
\begin{figure}
\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/preprocessed_highnoise}
\caption{A comparison of the preprocessed signals for the high noise case.}
\label{fig:PreprocessedHighNoise}
\end{figure}
\begin{figure}
\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/comparison_highnoise}
\caption{A comparison of the fitted signals for the high noise case.}
\label{fig:FitComparisonHighNoise}
\end{figure}
\begin{figure}
\includegraphics[clip=true,trim=6cm 2cm 6cm 3.5cm,width=17cm]{images/highnoise_56_noise}
\caption{Two particular preprocessed noise realizations for the high noise case.}
\label{fig:NoiseComparisonJustTwo}
\end{figure}

@@

\begin{figure}
\subfigure[$\tau_0$, $\alpha$, $E_0$, $V_0$, Run 1]
{\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=14cm]{images/highnoise_run5_1}}\\
\end{figure}
\begin{figure}
\subfigure[$\tau_s$, $\tau_f$, $\epsilon$, $V$, Run 1] 
{\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=14cm]{images/highnoise_run5_2}}\\
\end{figure}
\begin{figure}
\subfigure[$Q$, $S$, $F$, $BOLD$, Run 1 ]
{\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=14cm]{images/highnoise_run5_3}}
\caption{Converging histogram for parameters during run 1, as in \autoref{fig:NoiseComparisonJustTwo}.}
\label{fig:ConvergenceRuns1}
\end{figure}

\begin{figure}
\subfigure[$\tau_0$, $\alpha$, $E_0$, $V_0$, Run 2]
{\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=14cm]{images/highnoise_run6_1}}\\
\end{figure}
\begin{figure}
\subfigure[$\tau_s$, $\tau_f$, $\epsilon$, $V$, Run 2] 
{\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=14cm]{images/highnoise_run6_2}}\\
\end{figure}
\begin{figure}
\subfigure[$Q$, $S$, $F$, $BOLD$, Run 2 ]
{\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, width=14cm]{images/highnoise_run6_3}}
\caption{Converging histogram for parameters during run 2, as in \autoref{fig:NoiseComparisonJustTwo}.}
\label{fig:ConvergenceRuns2}
\end{figure}

@@

%NO SIGNAL, LOW NOISE
\subsection{Pure-Noise, low magnitude}
\label{sec:PureNoiseLowMag}
The two final single-voxel tests force the particle filter to attempt to "learn" a noise-only
time series. In the first test the noise used will be the same as that from the \autoref{sec:HighNoise},
$\sigma_x = .01, \sigma_y = .005$. The parameters will be set the same as well, but the
stimulus/input will be set to 0 the entire time. In effect it is a region of the brain for
which no visual stimulus induces activity. For the particle filter, on the other hand,
the stimulus will be left as in the previous two sections. The question then is how the
particle filter will respond to signal that does not correlate with the input, and how
the output may be differentiated from a time series that does. The time series'
are shown in \autoref{fig:NoiseOnly}, the preprocessed versions are shown in \autoref{fig:PreprocessedNoiseOnly}.
The line fits are shown in \autoref{fig:fits_noiseonly}.

\begin{figure}[H]
\centering
\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm,height=9cm]{images/realization_noiseonly}
\caption{Time-series lacking any real signal. With, $\sigma_x = .01, \sigma_y=.005$}
\label{fig:NoiseOnly}
\end{figure}

\begin{figure}[H]
\centering
\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm,height=9cm]{images/preprocessed_noiseonly}
\caption{A comparison of the preprocessed signals for the signal-free case. (
$\sigma_y = .01, \sigma_x = .005$)}
\label{fig:PreprocessedNoiseOnly}
\end{figure}

\begin{table}[t]
\centering
\begin{tabular}{|c | c | c | c | c | c | c | c | c |}
\hline 
$\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ & $\epsilon$  &  $\sqrt{MSR}$   \\
\hline 
1.0324 & 0.33211 & 0.34058 & 0.03012 & 1.40665 & 2.52079 & 0.5311 & \\
 0.98189 & 0.33047 & 0.3386 & 0.03014 & 1.45707 & 2.47232 & 0.45049 & 0.00158551 \\
 1.0429 & 0.33224 & 0.34124 & 0.02946 & 1.4618 & 2.49245 & 0.43012 &  0.00165055 \\
 1.02054 & 0.3321 & 0.33484 & 0.02586 & 1.45848 & 2.48741 & 0.4193 &  0.00150604 \\
 1.0565 & 0.33405 & 0.33758 & 0.02791 & 1.43784 & 2.52545 & 0.47517 & 0.00151513 \\
 1.01867 & 0.33528 & 0.33918 & 0.02782 & 1.48345 & 2.49605 & 0.44209 &0.00155645 \\
 1.051 & 0.33038 & 0.33837 & 0.02985 & 1.47651 & 2.48621 & 0.42719 &  0.00158526 \\
 1.00281 & 0.32929 & 0.33988 & 0.0298 & 1.43519 & 2.49256 & 0.48899 & 0.00164299 \\
 1.00893 & 0.33273 & 0.33982 & 0.0289 & 1.42903 & 2.49754 & 0.45688 & 0.0016793  \\
 1.01289 & 0.33275 & 0.3376 & 0.02997 & 1.41188 & 2.49881 & 0.50628 & 0.00182622 \\
 1.10247 & 0.33371 & 0.3419 & 0.02939 & 1.43774 & 2.53384 & 0.44079 & 0.00195009 \\
\hline                                                                
1.03009 & 0.33228 & 0.33905 & 0.02902 & 1.44506 & 2.50031 & 0.46076 & 0.00165129 \\
\hline 
\end{tabular}
\caption{Estimated Parameters on 11 different runs with low noise and no signal present.}
\label{tab:NoiseOnlyResults} 
\end{table}

\begin{figure}[H]
\centering
\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm,height=9cm]{images/fits_noiseonly}
\caption{Fits to the non-active, low noise signal. Note that the line is thick because all
the fits overlap. This is all 11 fitted lines.}
\label{fig:fits_noiseonly}
\end{figure}

\begin{figure}[H]
\centering
\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm,height=9cm]{images/justnoise_fit_0}
\caption{Fit from a single particle filter run, with the noise input. }
\label{fig:justnoise_fit_0}
\end{figure} %uses allnoise/ALLNOISE-0-w0

\begin{figure}[H]
\centering
\subfigure[$\tau_0$, $\alpha$, $E_0$, $V_0$]
{\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, height=9cm]{images/justnoise_hist_1}}\\
\end{figure}

\begin{figure}[H]
\centering
\subfigure[$\tau_s$, $\tau_f$, $\epsilon$, $V$] 
{\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, height=9cm]{images/justnoise_hist_2}}\\
\end{figure}

\begin{figure}[H]
\centering
\subfigure[$Q$, $S$, $F$, $BOLD$ ]
{\includegraphics[clip=true,trim=6cm 2cm 6cm 3cm, height=9cm]{images/justnoise_hist_3}}
\caption{Converging histogram for parameters when the signal consists purely of low level noise. 
Same run as \autoref{fig:justnoise_fit_0}}
\label{fig:JustNoiseConvergence}
\end{figure}

From the results it is clear that the parameters have not taken any visible turns into
unrealistic territory. It appears individual parameters will not be a distinguishing factor
in determining the validity of the model. Additionally the raw square-root mean-squared-residual
also appears to be lower than the SMSR of the low noise simulation from earlier. 
The precision with which the convergence occurred is actually the best seen in any
simulation.

However, the reason for low error may in fact be simply the fact that the over all
signal is significantly lower than any other simulation seen in other simulations.
The peaks never even cross 1\% difference, with most staying even below .5\% 
(\autoref{fig:PreprocessedNoiseOnly}). Because the signal is generally staying
well within the range of $.005$, which is the standard deviation of the weighting function
its unlikely that any particles will have a substantially different wieght. Thus, the reason
the fits are in such agreement is that no particles gained increased weight, and so
no convergence occurred at all, as further shown by \autoref{fig:JustNoiseConvergence}. 
Thus we have now seen how the particle filter responds when the signal is completely
random and low level, however there is also some possibility that the signal will
be random, yet active. Thus one more test is necessary on
a signal with enough noise to actually be considered a signal.

%NO SIGNAL, VERY HIGH NOISE
\subsection{Pure Noise, High Magnitude}
\label{sec:PureNoiseHighMag}
To determine how the particle filter will respond to active, yet unrelated
portions of the brain, in this section random data will again be used, but
with even higher standard deviations. To simulate this case another
pure noise signal was generated using a $\sigma_x$ of $.1$ and a $\sigma_y$ of $.05$.

As before, the convergence all seems to follow a similar path, leaving almost no
variance in the estimated time series (\autoref{fig:fits_noiseonly_high}). Now, however
the peaks are comparable to peaks in a "real" timeseries. Interestingly the algorithm
suffered from almost constant particle deprevation, meaning that the heuristic
for rescuing the particle filter from particle deprivation, discuessed in
\autoref{sec:Resampling} in fact was working against the proper course of action. Of course
 the proper course of action in this case would be for the particle filter to fail,
 since no particle can really properly estimate a random sequence. When this mechanism
 was removed, all 11 runs stopped due to particle deprevation (all weights hit zero). 
 The problem with allowing particle deprivation to occur is that it can \emph{rarely}
 occur if the prior happens to not be dense enough in the area of the solution. Thus,
 there is an issue of false positives vs. false negatives. A middle ground, such as 
 allowing deprivation to occur if the recovery mechanism has already triggered once
 may be an acceptable compromise, however it would be difficult to strike the 
 right balance. Instead for the purposes here we depend on the square root MSR to 
 evaluate if the results were reasonable or not.

\begin{figure}[H]
\centering
\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm,height=9cm]{images/fits_noiseonly_high}
\caption{Fits to the non-active, high noise signal. Note that the line is thick because all
the fits overlap. This is all 11 fitted lines.}
\label{fig:fits_noiseonly_high}
\end{figure}

\begin{figure}[H]
\centering
\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm,height=9cm]{images/justbignoise_fit_0}
\caption{Fit from a single particle filter run, with the noise input. }
\label{fig:justbignoise_fit_0}
\end{figure} %uses allnoise/ALLNOISE-0-w0

\begin{figure}[H]
\centering
\subfigure[$\tau_0$, $\alpha$, $E_0$, $V_0$]
{\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm, height=9cm]{images/justbignoise_1}}\\
\end{figure}
\begin{figure}[H]
\subfigure[$\tau_s$, $\tau_f$, $\epsilon$, $V$] 
{\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm, height=9cm]{images/justbignoise_2}}\\
\end{figure}
\begin{figure}[H]
\subfigure[$Q$, $S$, $F$, $BOLD$ ]
{\includegraphics[clip=true,trim=6cm 3cm 6cm 3cm, height=9cm]{images/justbignoise_3}}
\caption{Converging histogram for parameters when the signal consists purely of noise, with peaks comparable  to 
the convential BOLD signal. ($\sigma_y = .01, \sigma_x = .005$). Same run as fit in \autoref{fig:justbignoise_fit_0}}
\label{fig:JustNoiseConvergenceLarge}
\end{figure}

From \autoref{fig:justbignoise_fit_0} it is clear that, despite the fact that the input 
is competely detached from the signal, a non-zero response by the particle filter does
lead to decreased MSR. This is primarily due to the preprocessing which forces the signal to
have a non-zero mean. While this scheme does create a type of artificial activation for
the particle filter, in reality the MSR will still be less than if the signal was left
with zero mean, and the particle filter stayed flat. At best the particle filter will 
over-smooth creating a line that does down the center of the noisy signal. In such a case
the MSR will be the same as the alternative.
